---
title: "Simulation - Estimation of a full Langevin model"
author: "Pierre Gloaguen, Marie-Pierre Etienne"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
bibliography: 
  - biblio.bib
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
rm(list = ls())
knitr::opts_chunk$set(
  message = F, error = F, warning = F,
  comment = NA
)
```

```{r loadingPackage}
library(Rhabit)
library(viridis)
library(ggplot2)
library(gridExtra)
library(abind)
library(dplyr)
```


This vignette illustrates the use of the Rhabit package to simulate data from a Langevin model and to estimate the parameters.


# Simulation of the Langevin model

In the Langevin model, we assume that the movement is directed by a field of covariates. Those fields of covariates are mandatory to simulate a path according to the model.



## Simulating covariate fields

In the presented example, the path is observed on a square, centered on the origine and whose sides equal 30. 

In order to g a enerate a field of covariates, we propose to simulate two Mattern processes with the same parameters parameters $(\rho, \nu, \sigma^2)$  The covriates are obtained by applying a log transformation to the square of the Mattern process. 


```{r simCovParams}
set.seed(1)# repeatability
lim <- c(-15, 15, -15, 15) # limits of map
resol <- 0.1 # grid resolution
rho <- 4; nu <- 1.5; sigma2 <- 10# Matern covariance parameters
mean_function <- function(z){# mean function
  -log(3 + sum(z^2))
}
J <- 2 # number of covariates
```

Then, simulating

```{r simCovariates, eval = F}
covariates <- replicate(J, 
                        simSpatialCov(lim, nu, rho, sigma2, resol = resol,
                                      mean_function = mean_function,
                                      raster_like = T),
                        simplify = F)
```



The resulting covariates are available in the package~:

```{r loadingCovariates, echo = T, message=FALSE}
data("covariates", package = "Rhabit")
```

## Representation of the fileds of covariates 


```{r plot}
ggopts <- theme_light()+theme(text = element_text(size = 20),
                axis.text = element_blank(),
                axis.title = element_blank(),
                legend.key.height= unit(3, "line"),
                strip.background = element_rect(fill="white",colour = "grey") ,
                strip.text = element_text(size = 25, color = "black"))
levels <- factor(paste("Covariate", 1:J), levels = paste("Covariate", 1:J))
cov_df <- do.call(rbind.data.frame,
                 lapply(1:J,
                        function(j){
                          Rhabit::rasterToDataFrame(covariates[[j]], levels[j])
                        }))
ggplot(cov_df, aes(x,y)) + geom_raster(aes(fill = val)) +
  coord_equal() + scale_fill_viridis(name = "Value") + facet_wrap(level~.) +
  ggopts
```


## The resulting Utilization Distribution 

<!-- As detailed in the paper [@mgeb19] -->

\begin{equation}
	\rmd\X_t = \dfrac{\gamma^2}{2} \nabla \log \ud(\X_t) \rmd t + \gamma \rmd \bm{W}_t,\quad \X_0 =\bm{x}_0.
\end{equation}
```{r UD}
beta_true <- c(10, 5)
ud_rast <- covariates[[1]] # initialization
dx <- diff(ud_rast$x)[1]
dy <- diff(ud_rast$x)[1]
ud_rast$z <- exp(Reduce("+",
                    lapply(1:J, function(j)
                      dx * dy * beta_true[j] * covariates[[j]]$z)))
ud_rast$z <- ud_rast$z / sum(ud_rast$z)
ud_df <- rasterToDataFrame(ud_rast)
ggplot(ud_df, aes(x,y)) + geom_raster(aes(fill = val)) +
  coord_equal() + scale_fill_viridis(name = "Value") +
  ggopts
```


# Simulation

Setting simulation parameters:

```{r simLangevinMM}
t_final <- 10
n_obs <- 10001
times <- seq(0, t_final, length.out = n_obs)
```

Then, one can simulate using cov_list

```{r simulTracks, eval = F}
set.seed(1) #repeatability
tracks <- simLangevinMM(beta = beta_true, gamma2 = 10,
                        times = times, loc0 = c(0, 0),
                        cov_list = covariates, keep_grad = T)
```

These simulated tracks are a data set that can be loaded

```{r loadTracks}
data("tracks", package = "Rhabit")
```

And can be plotted

```{r}
ggplot(ud_df, aes(x,y)) + geom_raster(aes(fill = val)) +
  coord_equal() + scale_fill_viridis(name = "Value") +
  geom_path(data = tracks[, 1:2]) +
  ggopts
```

# Estimation

The dataset is expected to contain columns
-  `x` and `y` which specify the relocations
- `t` , for the corresponding times
- for each variable of interest, let's say  `V1`, two columns `V1_x` and `V1_y` which contain the computed gradient of covariates V1 at the recorded location.


```{r model_fitting}
## test with the formula
fitted_langevin <- fit_langevin_ud(
  cbind(x,y) ~ grad_c1 + grad_c2, data = tracks)
```

The effects of the covariates  can be obtained via 

```{r}
coef(fitted_langevin)
```


while the speed parameter is obtained through
```{r}
speed_coef(fitted_langevin)
```

# Testing thhe effects 

The function \code{summary} tests for every covariate, if the corresponding parameter equals 0 or not.

The Null hypothesis for the speed parameter is $\gamma^2 =0$.

```{r}
summary(fitted_langevin)
```


The confidence intervals for each parameter are available with.
```{r}
fitted_langevin$CI
```


The expected trend in the movement at each observed location is given by 

```{r}
head(fitted_langevin$predicted, n = 10)
```

